Resonant Scattering in a Uniform Sphere with Large Optical Depth

### 2021-11-19 ms.tex
A few last corrections & equation checks

A better approximation to the true solution has been derived by \citet{2006ApJ...649...14D}, who generalized the closed-form solution in slab geometry found in \citet{1990ApJ...350..216N}. It satisfies a $J=0$ boundary condition at $r=R$. Again, we generalize their solution to allow emission at frequency $\sigma_{\rm s}$ away from line center. The result can be written as a sum over spatial modes,
\be \label{eq:J0_sum}
J_0 = \frac{\sqrt{6}L}{16\pi \Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\frac{\sin{\kappa_n r}}{\kappa_n r}\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
and
\be \label{eq:H0_sum}
H_0 = - \frac{1}{3k\phi} \frac{\partial J_0}{\partial r} = -\frac{1}{3k\phi}\frac{\sqrt{6}L}{16\pi\Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\left(\frac{\cos{\kappa_n r}}{r} - \frac{\sin{\kappa_n r}}{\kappa_n r^2}\right)\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
which can be summed to give the closed form expressions
\be
J_0 & = & \frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right)
\label{eq:J0}
\ee
and
\be
H_0 & = &\frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr^2}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right. \nonumber \\ & & \left. - \left( \frac{\pi r}{R} \right)
\frac{ \cos(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
+ \left( \frac{\pi r}{R} \right)
\frac{ \sin^2(\pi r/R) }{ \left[ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R) \right]^2 }
\right).
\label{eq:H0}
\ee
These solutions agree with Equations \ref{eq:Jd} and \ref{eq:Hd} when the arguments of the trigonometric and hyperbolic functions are small. Again $J_0 \gg H_0$, except near $r=R$, where it goes to zero. The flux at $r=R$ can be written
\be
\nonumber
H_0(R, \sigma) & = & - \frac{1}{3k\phi}
\frac{\sqrt{6}L}{16\pi \Delta}
\frac{1}{R^3}
\sum_{n=1}^{\infty} 
n (-1)^n \exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)}\\
& = &  \frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi \Delta}
\frac{1}{R^3}
\left( 
\frac{ 1 }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] +1 }
\right).
\label{eq:H0surf}
\ee
Eq. \ref{eq:H0surf} will be shown to be a better approximation to the solution than Eq. \ref{eq:Hd}. It is still valid near the delta function at $r=0$, but is also a better approximation at $r=R$. $J_0$ decreases exponentially rather than as a power-law in frequency as it does for $J_d$, giving a much smaller flux on the line wings as compared to the divergent solution.

@@

@@

\begin{figure}
    \centering
    \includegraphics{final_residual.pdf}
    \caption{Spectrum $P(x)$ vs. frequency $x$. The analytic solutions for the outgoing spectrum are calculated at a line center optical depth of $\tau_0 = 1 \times 10^7$ and $a = 4.72\times 10^{-4}$ with $x_s = 0.0$. The top panel is linear scale, the middle panel is log scale with $|\rm H_{bc}|$ shown instead of $\rm H_{bc}$, and the bottom panel is the residual of each solution with Monte Carlo.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

@@

In Figure \ref{fig:sol_mc_residual_0}, the Monte Carlo escape probability as a function of frequency is shown along with that of the solutions $H_{\rm d}$, $H_0$, and $H_{\rm 0 + bc} = H_0 + H_{\rm bc}$ for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$.  The $H_{\rm bc}$ term is negative at the peak of the spectrum and positive in the line wing such that, when added to $H_0$, it corrects for the apparent excess of flux in the peaks of the spectrum. The solution with the correct frequency-dependent boundary condition enforced, $H_{0 + \rm bc}$, has lower residuals to Monte Carlo results than the other solutions, especially in the line wing. The boundary term corrects the deficit of $H_0$ in the line wings, further improving agreement with the numerical result. Note that the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term, since the Monte Carlo represents the ``true'' solution, $H$, and $H_{\rm bc} = H - H_0$. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings.

 \begin{figure}
    \centering
    \includegraphics{tau_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\tau_0 = 10^5, 10^6, 10^7$. The $x$ and $y$ axes are scaled by $(a\tau_0)^{1/3}$. Photons are initialized with $\rm x_s = 0.0$ in each case.}
    \label{fig:sol_mc_tau}
\end{figure}

@@

 \begin{figure}
    \centering
    \includegraphics{xinit_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\rm x_s = 0.0, 6.0, 12.0$. The optical depth at each of these source frequencies is $\tau_s = 10^7, 77,$ and $19$, respectively.} 
    \label{fig:sol_mc_xinit}
\end{figure}

 \begin{figure}
    \centering
    \includegraphics{xinit_threepanel_tau1e6.pdf}
    \caption{The same as Fig. \ref{fig:sol_mc_xinit}, but at a lower optical depth $\tau_0 = 10^6$. The shift $x_s$ is a much larger fraction of the distance to the spectral peak $(a\tau_0)^{1/3}$, and thus the asymmetry in the spectrum is much larger. The optical depth at the source frequency is $\tau_s = 10^6$, $7.7$, and $1.9$ for $x_s=0.0, 6.0,$ and $12.0$, respectively. } 
    \label{fig:sol_mc_xinit_lowtau}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit.pdf}
    \caption{$P_{\rm bc}(x) = 16\pi^2R^2H_{\rm bc}(R, x)\Delta/L$ (Eq. \ref{eq:prob_spectrum} with $H \rightarrow H_{\rm bc}$) vs. frequency $x$. Thicker lines indicate larger optical depth labeled by $\tau_0=10^5, 10^7,$ and $10^9$, and darker lines indicate $\sigma_s$ further from line center, labeled by $\sigma_s=0.0, \tau_0, 2\tau_0$. The location of the core-wing boundary $x_{cw}$ is shown with green vertical lines of increasing thickness for each $\tau_0$. The axes are scaled to show that the size of the correction factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

@@

Let us define the damping rate to be $\gamma \equiv i\omega$, which is real and positive for damped solutions. At the eigenvalues $\gamma=\gamma_{nm}$, the response $J(n, \sigma, \omega)$ is resonant. Near these $\gamma_{nm}$ poles, the eigenfunctions will have the form 
\be \label{eq:jnsigmaomega_approx}
J(n,\sigma,-i\gamma) & \simeq \frac{ J_{nm}(\sigma) }{\gamma_{nm} - \gamma} + C(\gamma, \sigma),
\ee
where $C(\gamma, \sigma)$ varies slowly in $\gamma$. If the $\omega$-integral in Equation \ref{eq:jrsigmat_expansion} could be closed at infinity and evaluated using the residue theorem, the result would be
\be
J(r,\sigma,t) & \simeq & j_0(\kappa_n r) J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Summing over all spatial modes $n$ and over all eigenmodes $m$ for a given $n$, we obtain
\be \label{eq:Jrsigmat}
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_{m=1}^{\infty} J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
This ansatz captures the contributions from $n \times m$ simple poles. Taking a derivative with respect to $r$ and evaluating at the surface $r=R$, we use
\be
\frac{dj_0(\kappa_n r)}{dr} \bigg\rvert_R & =& \frac{d}{dr} \left[ \frac{\sin(\kappa_n r)}{\kappa_n r} \right]\bigg\rvert_R
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)
\nonumber \\ & = & \frac{(-1)^n}{R}
\ee
to obtain the flux, which is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency emerging from the sphere to be
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating over time yields a factor $1/\gamma_{nm}$, and by integrating over $d\nu$ we find
\be \label{eq:sum_rule}
E & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3k} \sum_{nm} (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial ``sum rule'' provides a check on the values of $\gamma_{nm}$ and $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each mode is
\be \label{eq:pnmsoln}
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients are negative for odd values of $n$ and positive for even $n$. The size of each contribution scales roughly as $0.5/(m-7/8)^{2/3}$, with a weak dependence on $n$. This indicates the need for a large number of $n$ and $m$ to converge, in that it takes roughly ten times as many $m$ modes to reduce the size of $P_{nm}$ by a factor of ${\sim}$5.

@@

The eigenfunction's oscillatory forms have varying amplitudes which sum to create the final form of the mean intensity. The largest contribution at late times always comes from the ($n=1, m=1$) lowest-order eigenfunction. Fig. \ref{fig:jsoln} shows a set of eigenfunctions $J_{nm}(\sigma)$ to illustrate their relative scales for different $m$ at a fixed spatial eigenmode $n$. The overall scale of the $J_{nm}(\sigma)$ are set by the factor $E/(kR^3)$ with $E$ arbitrarily set to 1. For H atoms with $T=10^4$ K and $\tau_0=10^7$, an eigenfunction has typical size ${\sim} E a / \left(R^2 \Delta \right) = 10^{-37}$ in units of specific mean intensity times time. Additional terms add smaller-magnitude, faster-oscillating components that lead to higher accuracy upon summation with the lower-order terms. The oscillations of various modes must cancel, so many modes $m$ and $n$ are required for convergence to the correct solution. 
\begin{figure}
    \centering
    \includegraphics[width=\textwidth]{gamma_nm.pdf}
    \caption{Decay rate vs. mode number. Overlapping solid lines are plotted for each $n$. The thickness of the lines decrease with $n$ in order to clearly show where they are in agreement. These numerically-obtained resonant frequencies are compared with the analytic expression in Eq. \ref{eq:gamma_nm}. The parameters for the calculation are $\tau_0=10^7$ and $x_s=0.0$.}
    \label{fig:gamma_nm}
\end{figure}
The values of the $\gamma_{nm}$ can be described approximately with Eq. \ref{eq:gamma_nm}. Their values depend on $m$, $n$, and other physical parameters according to 
\be \label{eq:gamma_nm}
\gamma_{nm} = 2^{-1/3} \pi^{13/6} n^{4/3}\left(m-\frac{7}{8}\right)^{2/3}\frac{c}{R}(a\tau_0)^{-1/3}
\ee
as shown in Appendix \ref{app:wkb}. The power law in $m$ is weak, requiring up to $m=500$ to reduce the scale of $\gamma_{nm}^{\ \ -1}$ by two orders of magnitude. When sweeping through to find resonances, Equation \ref{eq:gamma_nm} is used to set the scale of the sweep points $\gamma_j$ to ensure no $\gamma_{nm}$ are missed. The continuity and close agreement with the analytic expression shown in Fig. \ref{fig:gamma_nm} indicate that the numerical solutions are accurate.

@@

\begin{figure}
    \centering
    \includegraphics{steadystate.pdf}
    \caption{Time-integrated spectrum $P(x)$ vs. $x$. The steady-state and time-integrated spectra for $n=1, ..., 20$ and $m=1, ..., 500$ are shown with $x_s=0$, $a = 4.72\times 10^{-4}$ and $\tau_0=10^7$.}
    \label{fig:steadystate}
\end{figure}

>>>>

A better approximation to the true solution has been derived by \citet{2006ApJ...649...14D}, who generalized the closed-form solution in slab geometry found in \citet{1990ApJ...350..216N}. It satisfies a $J=0$ boundary condition at $r=R$. Again, we generalize their solution to allow emission at frequency $\sigma_{\rm s}$ away from line center. The result can be written as a sum over spatial modes,
\be \label{eq:J0_sum}
J_0 = \frac{\sqrt{6}L}{16\pi \Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\frac{\sin{\kappa_n r}}{\kappa_n r}\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
and
\be \label{eq:H0_sum}
H_0 = - \frac{1}{3k\phi} \frac{\partial J_0}{\partial r} = -\frac{1}{3k\phi}\frac{\sqrt{6}L}{16\pi\Delta} \frac{1}{R^2}\sum_{n=1}^{\infty}n\left(\frac{\cos{\kappa_n r}}{r} - \frac{\sin{\kappa_n r}}{\kappa_n r^2}\right)\exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)},
\ee
where $\kappa_n=n\pi/R$. These can be summed to give the closed form expressions
\be
J_0 & = & \frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right)
\label{eq:J0}
\ee
and
\be
H_0 & = &\frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi^2 \Delta}
\frac{1}{Rr^2}
\left( 
\frac{ \sin(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
\right. \nonumber \\ & & \left. - \left( \frac{\pi r}{R} \right)
\frac{ \cos(\pi r/R) }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R)}
+ \left( \frac{\pi r}{R} \right)
\frac{ \sin^2(\pi r/R) }{ \left[ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] - \cos(\pi r/R) \right]^2 }
\right).
\label{eq:H0}
\ee
These solutions agree with Equations \ref{eq:Jd} and \ref{eq:Hd} when the arguments of the trigonometric and hyperbolic functions are small. Again $J_0 \gg H_0$, except near $r=R$, where it goes to zero. The flux at $r=R$ can be written
\be
\nonumber
H_0(R, \sigma) & = & - \frac{1}{3k\phi}
\frac{\sqrt{6}L}{16\pi \Delta}
\frac{1}{R^3}
\sum_{n=1}^{\infty} 
n (-1)^n \exp{\left(\frac{-\kappa_n \Delta}{k}|\sigma - \sigma_s|\right)}\\
& = &  \frac{1}{3k\phi}
\frac{\sqrt{6}L}{32\pi \Delta}
\frac{1}{R^3}
\left( 
\frac{ 1 }{ \cosh \left[ \frac{\pi \Delta}{k R} (\sigma - \sigma_s) \right] +1 }
\right).
\label{eq:H0surf}
\ee
Eq. \ref{eq:H0surf} will be shown to be a better approximation to the solution than Eq. \ref{eq:Hd}. It is still valid near the delta function at $r=0$, but is also a better approximation at $r=R$. $J_0$ decreases exponentially rather than as a power-law in frequency as it does for $J_d$, giving a much smaller flux on the line wings as compared to the divergent solution.

@@

\begin{figure}
    \centering
    \includegraphics{final_residual.pdf}
    \caption{Spectrum $P(x)$ vs. frequency $x$. The analytic solutions for the outgoing spectrum are calculated at a line center optical depth of $\tau_0 = 1 \times 10^7$ with $x_s = 0.0$. The top panel is linear scale, the middle panel is log scale with $|\rm H_{bc}|$ shown instead of $\rm H_{bc}$, and the bottom panel is the residual of each solution with Monte Carlo.} 
    \label{fig:sol_mc_residual_0}
\end{figure}

@@

@@

In Figure \ref{fig:sol_mc_residual_0}, the Monte Carlo escape probability as a function of frequency is shown along with that of the solutions $H_{\rm d}$, $H_0$, and $H_0 + H_{\rm bc}$ for an optical depth of $\tau_0 = 10^7$ and photons emitted at line center $\rm x_s = 0$.  The $H_{\rm bc}$ term is negative at the peak of the spectrum and positive in the line wing such that, when added to $H_0$, it corrects for the apparent excess of flux in the peaks of the spectrum. The solution with the correct frequency-dependent boundary condition enforced, $H_0 + H_{\rm bc}$, has lower residuals to Monte Carlo results than the other solutions, especially in the line wing. The boundary term corrects the deficit of $H_0$ in the line wings, further improving agreement with the numerical result. Note that the residuals to the $H_0$ solution are a close match to the $H_{\rm bc}$ term, since the Monte Carlo represents the ``true'' solution, $H$, and $H_{\rm bc} = H - H_0$. It is evident that the divergent solution $H_{\rm d}$ fails in the line wings.

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{tau_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\tau_0 = 10^5, 10^6, 10^7$. The $x$ and $y$ axes are scaled by $(a\tau_0)^{1/3}$. Photons are initialized with $\rm x_s = 0.0$ in each case.}
    \label{fig:sol_mc_tau}
\end{figure}

@@

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit_threepanel.pdf}
    \caption{The same as Figure \ref{fig:sol_mc_residual_0}, but for $\rm x_s = 0.0, 6.0, 12.0$. The optical depth at each of these source frequencies is $\tau_s = 10^7, 77,$ and $19$, respectively.} 
    \label{fig:sol_mc_xinit}
\end{figure}

 \begin{figure}
    \centering
    \includegraphics[width=\textwidth]{xinit_threepanel_tau1e6.pdf}
    \caption{The same as Fig. \ref{fig:sol_mc_xinit}, but at a lower optical depth $\tau_0 = 10^6$. The shift $x_s$ is a much larger fraction of the distance to the spectral peak $(a\tau_0)^{1/3}$, and thus the asymmetry in the spectrum is much larger. The optical depth at the source frequency is $\tau_s = 10^6$, $7.7$, and $1.9$ for $x_s=0.0, 6.0,$ and $12.0$, respectively. } 
    \label{fig:sol_mc_xinit_lowtau}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{xinit.pdf}
    \caption{$P_{\rm bc}(x) = 16\pi^2R^2H_{\rm bc}(R, x)\Delta/L$ (Eq. \ref{eq:prob_spectrum} with $H \rightarrow H_{\rm bc}$) vs. frequency $x$. Thicker lines indicate larger optical depth labeled by $\tau_0=10^5, 10^7,$ and $10^9$, and darker lines indicate $\sigma_s$ further from line center, labeled by $\sigma_s=0.0, \tau_0, 2\tau_0$. The location of the core-wing boundary $x_{cw}$ is shown with green vertical lines of increasing thickness for each $\tau_0$. The axes are scaled to show that the size of the correction factor agrees with the predicted $(a\tau_0)^{-1/3}$ scaling.}
    \label{fig:xinit}
\end{figure}

@@

Let us define the damping rate to be $\gamma \equiv i\omega$, which is real and positive for damped solutions. At the eigenvalues $\gamma=\gamma_{nm}$, the response $J(n, \sigma, \omega)$ is resonant. Near these $\gamma_{nm}$ poles, the eigenfunctions will have the form 
\be \label{eq:jnsigmaomega_approx}
J(n,\sigma,-i\gamma) & \simeq \frac{ J_{nm}(\sigma) }{\gamma_{nm} - \gamma} + C(\gamma, \sigma),
\ee
where $C(\gamma, \sigma)$ varies slowly in $\gamma$. If the $\omega$-integral in Equation \ref{eq:jrsigmat_expansion} could be closed at infinity and evaluated using the residue theorem, the result would be
\be
J(r,\sigma,t) & \simeq & j_0(\kappa_n r) J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Summing over all spatial modes $n$ and over all eigenmodes $m$ for a given $n$, we obtain
\be \label{eq:Jrsigmat}
J(r,\sigma,t) & = & \sum_{n=1}^\infty j_0(\kappa_n r)  \sum_{m=1}^{\infty} J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
This ansatz captures the contributions from $n \times m$ simple poles. Taking a derivative with respect to $r$ and evaluating at the surface $r=R$, we use
\be
\frac{dj_0(\kappa_n r)}{dr} \bigg\rvert_R & =& \frac{d}{dr} \left[ \frac{\sin(\kappa_n r)}{\kappa_n r} \right]\bigg\rvert_R
=  \left( \frac{\cos(\kappa_n R)}{R} - \frac{\sin(\kappa_n R)}{\kappa_n R^2} \right)
\nonumber \\ & = & \frac{(-1)^n}{R}
\ee
to obtain the flux, which is
\be
F(R,\sigma,t) & =& - \frac{4\pi}{3k\phi} \frac{dJ(R,\sigma,t)}{dr} 
= - \frac{4\pi}{3k\phi R}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\ee
Multiplying by $4\pi R^2$ gives the energy per time per frequency emerging from the sphere to be
\be
\frac{dE}{dtd\nu} & = & - \frac{16\pi^2 R}{3k\phi}  \sum_{nm} (-1)^n J_{nm}(\sigma) e^{-\gamma_{nm}t}.
\label{eq:dEdtdnu}
\ee
Integrating over time yields a factor $-1/\gamma_{nm}$, and by integrating over $d\nu$ we find
\be \label{eq:sum_rule}
E & = &  \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3k} \sum_{nm} (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
This non-trivial ``sum rule'' provides a check on the values of $\gamma_{nm}$ and $J_{nm}(\sigma)$. This expression can also be written as
\be
1 & =& \sum_{nm} P_{nm},
\label{eq:sumrule}
\ee
where the contribution of each mode is
\be \label{eq:pnmsoln}
P_{nm} & \equiv & \sqrt{ \frac{3}{2} } \frac{16\pi^2R\Delta^2}{3kE}  (-1)^{n+1} \gamma_{nm}^{-1} \int d\sigma J_{nm}(\sigma).
\ee
These coefficients are negative for odd values of $n$ and positive for even $n$. The size of each contribution scales roughly as $0.5/(m-7/8)^{2/3}$, with a weak dependence on $n$. This indicates the need for a large number of $n$ and $m$ to converge, in that it takes roughly ten times as many $m$ modes to reduce the size of $P_{nm}$ by a factor of ${\sim}$5.

@@

The eigenfunction's oscillatory forms have varying amplitudes which sum to create the final form of the mean intensity. The largest contribution at late times always comes from the ($n=1, m=1$) lowest-order eigenfunction. Fig. \ref{fig:jsoln} shows a set of eigenfunctions $J_{nm}(\sigma)$ to illustrate their relative scales for different $m$ at a fixed spatial eigenmode $n$. The overall scale of the $J_{nm}(\sigma)$ are set by the factor $E/(kR^3)$ with $E$ arbitrarily set to 1. For H atoms with $T=10^4$ K and $\tau_0=10^7$, an eigenfunction has typical size ${\sim} E a / \left(R^2 \Delta \right) = 10^{-37}$ in units of specific mean intensity times time. Additional terms add smaller-magnitude, faster-oscillating components that lead to higher accuracy upon summation with the lower-order terms. The oscillations of various modes must cancel, so many modes $m$ and $n$ are required for convergence to the correct solution. 
\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{gamma_nm.pdf}
    \caption{Decay rate vs. mode number. Overlapping solid lines are plotted for each $n$. The thickness of the lines decrease with $n$ in order to clearly show where they are in agreement. These numerically-obtained resonant frequencies are compared with the analytic expression in Eq. \ref{eq:gamma_nm}. The parameters for the calculation are $\tau_0=10^7$ and $x_s=0.0$.}
    \label{fig:gamma_nm}
\end{figure}
The values of the $\gamma_{nm}$ can be described approximately with Eq. \ref{eq:gamma_nm}. Their values depend on $m$, $n$, and other physical parameters according to 
\be \label{eq:gamma_nm}
\gamma_{nm} = 2^{-1/3} \pi^{13/6} n^{4/3}\left(m-\frac{7}{8}\right)^{2/3}\frac{c}{R}(a\tau_0)^{-1/3}
\ee
as shown in Appendix \ref{app:wkb}. The power law in $m$ is weak, requiring up to $m=500$ to reduce the scale of $\gamma_{nm}^{\ \ -1}$ by two orders of magnitude. When sweeping through to find resonances, Equation \ref{eq:gamma_nm} is used to set the scale of the sweep points $\gamma_j$ to ensure no $\gamma_{nm}$ are missed. The continuity and close agreement with the analytic expression shown in Fig. \ref{fig:gamma_nm} indicate that the numerical solutions are accurate.

@@

\begin{figure}
    \centering
    \includegraphics{steadystate.pdf}
    \caption{Time-integrated spectrum $P(x)$ vs. $x$. The steady-state and time-integrated spectra for $n=1, ..., 20$ and $m=1, ..., 500$ are shown with $x_s=0$ and $\tau_0=10^7$.}
    \label{fig:steadystate}
\end{figure}
